---
title: "Exploratory plots on 2021 AO3 data dump"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(ggplot2)
```

## Import data

Before this file, I ran `data_import_save.R`, which does minimal cleaning on the original CSV files and then exports them into `.Rda` format (loads faster, takes less disk space). 

```{r}
load("data/works.Rda")
works_raw <- works
```

## Basic initial plots - Works

### Number of works posted by date

First, I was curious about the works posted over time, as a whole and then by restricted and completed status. There are 7269693 works listed in the file.

```{r works-hist, message=FALSE}
works %>% ggplot(mapping = aes(x = creat_date)) +
  geom_histogram()
```

By complete/incomplete, the numbers are:

```{r works-hist-complete, echo=FALSE, message=FALSE}
works %>% 
  count(complete) %>% 
  mutate(frac = n / nrow(works))

works %>% ggplot(mapping = aes(x = creat_date)) +
  geom_histogram() + 
  facet_grid(cols = vars(complete)) + 
  ggtitle("Works posted by date, by completion status")
```

By restricted/unrestricted (here, plotting fraction of works rather than raw number, or the restricted data gets totally squashed)

```{r works-hist-restrict, echo=FALSE, message=FALSE}
works %>% 
  count(complete) %>% 
  mutate(frac = n / nrow(works))

works %>% ggplot(mapping = aes(x = creat_date)) +
  geom_histogram(aes(y = ..density..)) + 
  facet_grid(cols = vars(restricted)) + 
  ggtitle("Works posted by date, by restricted status")
```

Slightly interesting that restricted works don't seem to increase at the same rate. Maybe the pandemic makes people less shy? Who knows. 


### Cumultative works over time

Next, I'm curious how the total evolved over time. This is easier if I start by accumulating the counts of each work type. 

```{r}
works <- works_raw %>% 
  arrange(desc(row_number())) %>%   # invert rows first to count up
  mutate(ctAll = row_number()) %>%  # count all works
  group_by(complete) %>%     # count complete works
  mutate(ctComplete = row_number()) %>% 
  ungroup() %>% 
  group_by(restricted) %>%   # now do restricted
  mutate(ctRestrict = row_number()) %>% 
  ungroup()

works
```

The cumulative plot of all works over time looks like this. Second version was to see how exponential it was (answer: meh). 

```{r works-cumulative}
ct_all <- works %>% 
  group_by(creat_date) %>% 
  summarize(ctAll = max(ctAll))   

# All works over time
ct_all %>% ggplot(aes(x = creat_date, y = ctAll)) +
  geom_line() + 
  ggtitle("Cumulative works over time")

# How exponential is it?
ct_all %>% ggplot(aes(x = creat_date, y = ctAll)) +
  geom_line() + 
  scale_y_log10() + 
  ggtitle("Cumulative works over time")
```

Then repeat, breaking out by complete/incomplete: 

```{r works-cumulative-complete, message=FALSE}
ct_complete <- works %>% 
  group_by(creat_date, complete) %>% 
  summarize(ctComplete = max(ctComplete))

# Complete works over time
ct_complete %>% 
  ggplot(mapping = aes(x = creat_date, y = ctComplete)) + 
  geom_line(aes(color = complete))  +
  ggtitle("Completed work count over time")
```

And restricted/unrestricted: 

```{r works-cumulative-restrict, message=FALSE}
ct_restrict <- works %>% 
  group_by(creat_date, restricted) %>% 
  summarize(ctRestrict = max(ctRestrict)) 

# Restricted works over time
ct_restrict %>% 
  ggplot(mapping = aes(x = creat_date, y = ctRestrict)) + 
  geom_line(aes(color = restricted)) +
  ggtitle("Restricted work count over time")
```


## Language frequency

Just plotting work frequency by language isn't super helpful, because English is too common. 

```{r language-freq}
works %>% 
  count(language) %>% 
  ggplot(aes(x = language, y = n)) + geom_col()
```

If you filter that away, the axis labels are still hard to read. There are 87 different languages represented! Here, I set an arbitrary threshold of at least 50 works to display.

```{r language-freq-noen}
# Axis labels hard to read without filtering for frequency
works %>% 
  count(language) %>% 
  filter(language != "en", n >= 50) %>% 
  ggplot(aes(x = n, y = language)) + 
  geom_col() +
  ggtitle("Frequency of works by language (omitting English)")
```

Next, I was curious about development over time. First, filter down to the most frequent languages, excluding English: 

```{r, message=FALSE}
top_lang <- works %>% 
  #filter(language != "en") %>% 
  count(language) %>% 
  slice_max(order_by = n, n = 10) %>% 
  mutate(frac = n / nrow(works))

top_lang

ct_lang <- works %>% 
  filter(language %in% top_lang$language) %>% 
  group_by(language) %>% 
  mutate(ctLang = row_number()) %>%  # count all works
  group_by(creat_date, language) %>% 
  summarize(ctLang = max(ctLang))
```

I cut off at 10 because it's a round number, but it also turns out there's a natural breakpoint--the next most frequent after Polish (14261 works) is Turkish, with substantially fewer (2984) works. 

```{r language-time}
ct_lang %>% 
  filter(language != "en") %>% 
  ggplot(aes(x = creat_date, y = ctLang)) +
  geom_line() + 
  facet_wrap(vars(language)) + 
  ggtitle("Works over time, top non-English languages")  
```

The really visible thing here is the big jump, ramping up a few years ago, in the frequency of Chinese-language (`zh`) works. To get a slightly better look at that and how it compares to the other growth rates, let's try one more: 

```{r language-time-compare2}
ct_lang %>% 
  filter(language != "en") %>% 
  ggplot(aes(x = creat_date, y = ctLang)) +
  geom_line(aes(color = language)) + 
  scale_y_log10() +
  ggtitle("Works over time, top non-English languages")  
```

The Chinese-language section has been growing faster than many other languages for a bit, starting around 2012 and with another change for the steeper around 2018-2019. 


## First look - Tags

Now pull in the tags data set and take a look at the most common ones. 

```{r}
load("data/tags.Rda")
tags_raw <- tags

tags %>% arrange(desc(cached_count))
```

Not a surprise that the most common ones are from big categories (archive warnings, ratings, relationship categories). I am amused that Fluff made the top list, though. Makes me wonder if it or any other tags became notably more popular in 2020. 


### Tag type frequency

Other than the most popular, there are some big categories that I don't really have a sense of the shape of: 

* How many tags are unique (vs. having a merger ID)
* How many are listed but not actually used
* What the overall frequency distribution looks like
* How much of the list is Redacted (low-frequency tags, whose names aren't listed)
* How common different rating tags and media types are

Going from the top, any tag row that has a `merger_id` is not a unique tag, but a different listing for an already-existing tag. For example:

```{r}
tags %>% 
  filter(!is.na(merger_id)) %>% 
  slice_head(n = 5)
```

Tag 29 (Sam Winchester/Dean Winchester) is merged to tag 35 (Dean Winchester/Sam Winchester), the canonical version. I think this might mean canonical tags won't have a `merger_id`? But I can check that. First, the overall prevalence of unique vs. merged tags: 

```{r}
tags %>% count(is.na(merger_id)) %>% 
  rename(unique = contains("na")) %>% 
  mutate(frac = n / nrow(tags))
```

About 24% of the tag database is duplicate entries--different spellings or wordings that link back to the same tag. As to my question about if `merger_id` implies that `canonical == FALSE`: 

```{r}
tags %>% 
  count(is.na(merger_id), canonical) %>% 
  rename(unique = contains("na")) %>% 
  arrange(desc(unique))
```

So... not quite, actually! There are 51 canonical tags that still point to a merger. Taking a quick look...

```{r}
tags %>% 
  filter(!is.na(merger_id), canonical == TRUE) %>% 
  slice_head(n = 5)
```

And then retrieving the first couple to see what they point to:

```{r}
tags %>% filter(id == 87022 | id == 514794)
```

At least in those two cases, both the "base" tag and the one with a `merger_id` are canonical. 

Anyway, moving along to my second question: How many are listed but not actually used?

```{r}
tags %>% 
  filter(cached_count == 0) %>% 
  tally() %>% 
  mutate(frac = n / nrow(tags))
```

Just under 10% of tags have a `cached_count` of zero. I'm not sure this is actually the same as never being used (see below), and will have to see what happens when I process the tags listing column of the `works` frame.

Next: What does the overall frequency distribution look like? (The answer is "one enormo spike" on the linear scale, so let's look at log-log.)

```{r tag-freq, warning=FALSE}
tags %>% ggplot(aes(x = cached_count)) + 
  geom_histogram(bins = 50) + 
  scale_y_log10() + 
  scale_x_log10() +
  ggtitle("Frequency of tag use")
```

This looks pretty power-law-ish. Most of the tags, more than ten million of them, are used only once (or at least have a `cached_count` of 1). At the other extreme, a handful of tags show up millions of times.


Next question: How much of the list is Redacted (low-frequency tags, whose names aren't listed)? There are at least two ways to ask this--how many of those tags exist, and how commonly they show up in works. Below, the first question is tabulated in `n`, and the second in `cached_count`. 

```{r}
tags %>% 
  mutate(redacted = (name == "Redacted")) %>%   # not sure I want to keep this
  group_by(redacted) %>% 
  add_count() %>% 
  summarize(cached_count = sum(cached_count), n = n()) %>% 
  mutate(cachefrac = cached_count / sum(cached_count),
         nfrac = n / nrow(tags))
```
The last two columns convert those numbers to a fraction of the total. So redacted (low-use) tags are prolific, almost 90% of the total tag list, which seems plausible if I think about how many tumblr-style one-off notes appear that way. But 90% of the cached count comes from non-redacted AKA higher-frequency tags. And there is a single tag with an `NA` for a name, and 708 uses. Huh. 


Next: How common are different rating tags and media types? Ratings first...

```{r}
tags %>% filter(type == "Rating") %>% 
  select(name, cached_count) %>% 
  mutate(frac = cached_count / sum(cached_count)) 

```

Mostly General and Teen, which surprised me a bit but probably shouldn't. So much conversation focuses around the explicit end of the rating scale, it's easy to forget that it's only about a sixth of the material. Now, looking at media types:

```{r}
tags %>% 
  filter(type == "Media")  # this is way fewer than I expected...

```

Here, something seems fishy. Either I'm misunderstanding what `cached_count` means, or I don't understand how the Media tag type gets applied, because I expected *much* higher counts. I'll need to come back to this once I'm combining the works and tags frames. 