---
title: "Workflow to merge works/tags info for a fandom"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
```

For most questions that interest me, I need to merge information from the tags and works databases. Here's a workflow for doing that, after a few hours and a short stint in regex purgatory. I'm using the example of pulling all info for a particular fandom, but it could just as easily be all Teen-rated works, everything associated with video games, or whatever.


## Import data

Before this file, I ran:

1. `data_import_save.R`, which does minimal cleaning on the original CSV files and then exports them into `.Rda` format. This includes reordering works to go from oldest to newest and adding an ID number column (`wid`).

2. `merge_works.R`, which replaces any duplicate tag with its "real" (`merger_id`) counterpart.

```{r}
(load("data/works_merged.Rda"))
(load("data/tags.Rda"))

works <- works_merged
rm(works_merged)
```

Peek at the top of each:

```{r}
works
tags
```


## Helper functions

Two tasks I'll need: Find the tag ID number associated with a given name in the `tags` frame, and find all rows containing a given tag ID in the `works` frame. 

```{r}
# Regex to match: Start of line OR non-digit, tag, non-digit OR end of line
makePat <- function(x) {
  pat <- paste0("(^|[^[:digit:]])", x, "([^[:digit:]]|$)") 
}

# Find tag number for a given name
getTagID <- function(x, df = tags) {
  df %>% filter(name == x) %>% pull(var = id)
}

getTagID("RWBY")
```


## Pick a fandom and filter `works` for it

At the end of this process, I'd like three outputs: 

1. The fandom-specific subset of the `works` frame, otherwise in its original format. 

2. A long data frame version of 1, with one tag per row. 

3. The tag list filtered down to contain only tags appearing in 1 and 2.


### Filter the `works` frame

Start by pulling the row(s) for a given tag.

```{r}
tagpat <- makePat(getTagID("RWBY"))

wred <- works %>% 
  filter(grepl(pattern = tagpat, x = tags))

wred
```

We're now down to about 26K works, a respectable but not overwhelming fraction of the data set. 


### Make a long `works` data frame

Next, unroll the tag column into a long data frame, by way of a list-column.

```{r}
wlong_fandom <- wred %>% 
  mutate(tag_list = stringr::str_split(.data$tags, 
                                       stringr::fixed("+"))) %>% 
  unnest(tag_list) %>% 
  mutate(tag_list = as.integer(tag_list)) %>% 
  select(-tags)

wlong_fandom
```

In the long data frame, I've dropped the original `tags` column, which holds the character string of tag IDs separated by + symbols. It's still there in the `wred` frame, but it's not useful and hurts readability to repeat it every row of `wlong`. 

Finally, for ease of reading I want to attach tag names to my long works frame. Might as well add tag type while I'm in the neighborhood. 

```{r}
wtagged_fandom <- wlong_fandom %>% 
  left_join(tags %>% select(id, type, name), 
            by = c("tag_list" = "id"))

wtagged_fandom
```

Check: Are any of these tags missing a match in the `tags` frame?

```{r}
wtagged_fandom %>% 
  mutate(mismatch = is.na(name)) %>% 
  group_by(mismatch) %>% 
  count() %>% 
  mutate(frac = n / nrow(wtagged_fandom))

# Inspect any rows without a tag match
wtagged_fandom %>% 
  mutate(mismatch = is.na(name)) %>% 
  filter(mismatch) 
```

Nine of my rows came up with `NA` for a tag name. Looking at the first of them, 54605328:

```{r}
tags %>% filter(id > 54605320)
```

There are a number of skipped `id` values, and this `NA` from my matched RWBY tags is one of them. To be really careful I'd check them all, but for the moment I'll assume those represent deleted tag entries. 


## Filter the `tags` frame to in-use entries

There are 14.5 million tags and many of them don't apply to this subset of works, so I also want to cut down the `tags` frame to its corresponding subset. 

After everything above, this goes pretty fast. Pull in-use tags from filtered works frame:

```{r}
tags_used <- wlong_fandom %>% 
  ungroup() %>%   # had a wid group, might want to check back for problems there
  select(tag_list) %>% 
  distinct() %>% 
  arrange(tag_list)

tags_used
```

Then join tags frame info to each in-use tag:

```{r}
tred <- tags_used %>% 
  left_join(tags, by = c("tag_list" = "id")) %>% 
  rename(id = tag_list)

tred
```

That's about 61K tags. When I did this the first time, without merging duplicate tags in the works list first, my filtered list of in-use tags was 74,026 entries long. So I've cut off more than 15% of my tags by cleaning up the list first. 

And we're done, unless I find a bug later.


## Save data

Even though the data frames are a much more manageable size now, I'll keep saving the tags and works info in separate files. For works, I'll keep both the subsetted-but-wide-format `wred`, and the one-row-per-tag `wtagged`. 

```{r}
#save(tred, file = "data/tags_RWBY.Rda")
#save(wred, wtagged_fandom, file = "data/works_RWBY.Rda")
```

